### 2022-03-15 thesis/introduction.Rmd

---
title: ''
output: pdf_document
bibliography: ../references.bib
---

# Introduction

## Background

<!-- VO2max is important -->
<!-- physiological determinants -->
<!-- VO2max testing -->
<!-- Influence factors in testing -->

## Previous Research

<!-- First example in 90s: applied research -->
<!-- Since then: various example -->
<!-- conceptual work: Robergs -->
<!-- Comparison: martin-rincon -->
<!-- Problem: group + individual -->
<!-- Knowledge Gap -->

## Aim

<!-- Gap -> we need this: -->
<!-- overview what is actually done -->
<!-- and how does this impact VO2max assessment -->

<!-- Review: mapping current methodological (+ reporting) state  -->
<!-- Comparison: highlight advantage over previous -->

<!-- help compare results across studies + for individuals -->
<!-- advise for use of methods  -->
<!-- call for improved reporting -->

@bassett2000; @tricco2018