### 2022-03-15 thesis/results.Rmd

---
title: ''
output: pdf_document
bibliography: ../references.bib
---

# Results

## Systematic Scoping Review

## Experimental Comparison

![A plot](../preregistration/example_plot.png)