### 2022-04-28 thesis/discussion.Rmd
Continue writing on thesis

---
title: ''
output: pdf_document
bibliography: ../references.bib
---

# Discussion

<!-- if report - diverse strategies -->
<!-- most (multiple) binned time averages: necessary for mixing chamber, but bbb has more information! -->
<!-- current practice is in contrast to recommendations (Robergs)! -->

<!-- variety of calculation intervals - different results (example)-->
<!-- comparing results for strategies -->

<!-- recommendation for choosing a strategy: ... -->

<!-- guidelines for reporting: ... -->