### 2022-04-28 thesis/results.Rmd

---
title: ''
output: pdf_document
bibliography: ../references.bib
---

# Results

## Systematic Scoping Review

Initial search yielded 7529 results of which 4364 remained after automated filtering and removal of duplicates (see flow diagram in Figure 1). Out of the random sample (n = 500), 244 articles were included in the final analysis.

![Flow diagram for the systematic scoping review in accordance with the PRISMA 2020 Statement [@page2021]](../plots/flow_diagram.png)

Reporting practices of the methodology of gas exchange measures differed widely within the literature (see Table 4). More than half (51.8%) of the articles did not report any information on their data processing strategy. Only one in twenty articles (5.4%) provided a rationale for their used strategy.

```{r reporting-table, echo=FALSE}
load("../data/reporting.Rda")
t <- purrr::map_dfr(reporting, scales::percent, accuracy = 0.1)

t[["pre"]] <- paste0(t[["pre"]], "*")
t[["software"]] <- paste0(t[["software"]], "*")

knitr::kable(
  t,
  col.names = c("Metabolic cart", "Preprocessing", "Software", "Processing Strategy", "Reference"),
  caption = "Percentage of studies that provided details on the different characteristics of oxygen uptake data processing. *only examined within the subgroup of studies using breath-by-breath measurements",
  align = "r"
)
```

Out of the authors that provided information and collected breath-by-breath measurements most (79.5%) utilized binned averages to determine V̇O~2max~. Moving time averages, or breath-based averages were uncommon (see Figure 2). No study used digital filtering methods to determine V̇O~2max~.

![Data strategies for processing breath-by-breath data in the reviewed literature (n = 88).](../plots/strategies.png)

The calculation interval for time-based averages of mixing chamber and breath-by-breath devices ranged from 5 to 60 seconds (see Figure 3). 30 second length intervals were most common to define V̇O~2max~, while authors also often employed shorter (10-20s) and longer (60s) periods.

![Total durations of the calculation interval of V̇O~2max~ in the reviewed studies.](../plots/duration_count.png)

## Experimental Comparison

![VO2max varies by data processing strategy. Values are expressed relative to the VO2max from a 30-second binned average --- the most common strategy as determined by the review. Solid lines display the median, the shaded area marks the interval between 10th and 90th percentile. Using moving average leads to systematically higher VO2max values compared to binned time averages. Changing the averaging interval (in seconds or breaths) can lead to changes in VO2max as large as 5%.](../plots/comparison.png)

![Respiratory Rates peak around 60min^-1^ in the ramp tests. The red segments correspond to the last minute before exhaustion of each individual (n = 72). .](../plots/rr.png)