### 2022-05-10 thesis/discussion.Rmd
Update thesis pdf

---
title: ''
output: pdf_document
bibliography: ../references.bib
---

# Discussion

The collected data shows, that current research uses a variety of strategies to determine V̇O~2max~, which directly influences the values obtained. Many articles only incompletely report on their methods and use outdated strategies. These practices hinder the validity and reproducibility of V̇O~2max~ measurement.

## Current state of data processing

Despite the call to use moving averages [@robergs2010], binned time averages remain the most commonly used in the reviewed literature.
<!-- compare with midgley? differences? why? highlight systematic aspect? -->
<!-- ...different journals, not self-assessed -->
<!-- explain why binned leads to lower vo2max - see fig 1 -->
<!-- Why do researchers use binned? Not aware of problems? -->
<!-- binned common for mixing (but not necessary!). Removes major advantages of bbb -->
<!-- not reporting at all - compare number -->

<!-- length of calculation interval: very important (example), but often differing -->
<!-- may be due to opposing recommendations -->
<!-- rational for both short and long, reporting even more important -->

<!-- project allows to compare data among strategies -->
<!-- but: very individual (and most research only on group mean) -->
<!-- : for comparison of processing strategies in individuals simply reporting is not enough -->

<!-- recommendation for choosing a strategy: digital filter! -->
<!-- but zero phase (not as robergs) -->
<!-- practical issue: requires advanced software -->

<!-- guidelines for reporting: report! -->
<!-- checklist -->

<!-- even better: share data and code (reporducible analysis) -->
<!-- data is easy to anonymize -->
<!-- tools have been recently developed (whippr, spiro) -->