### 2022-05-10 thesis/conclusion.qmd
Move thesis from R Markdown to Quarto

---
title: ''
format: pdf
bibliography: ../references.bib
---

# Conclusion