### 2022-05-17 thesis/introduction.qmd
Adjust to new citation style

---
title: ''
format: pdf
bibliography: ../references.bib
---

# Introduction

## Background

Performance in endurance sports is limited by the human physiology. Athletes need to supply energy to the contracting muscles for locomotion, a process that happens mainly via the oxidative phosphorylation. A higher maximal activity of the oxidative phosphorylation allows to supply more energy, and thus to move faster [@holloszy1984]. The highest activity of oxidative phosphorylation on a cellular level can be approximated by measuring the maximum oxygen uptake (V̇O~2max~) on a whole-body level.

V̇O~2max~ is defined as the highest rate a body can consume oxygen [@bassett2000]. Determined by measuring gas exchange data, it is one of the most common assessed exercise parameters in sports science and medicine. V̇O~2max~ highly corresponds with endurance performance in heterogeneous groups [@reaburn2008; @costill1973; @tanaka1990] and can be regarded the most relevant physiological parameter for predicting endurance performance [@bassett2000]. Many training programs aim to target the V̇O~2max~, and studies evaluate the quality of interventions by measuring changes in V̇O~2max.~

Researchers predominantly measure V̇O~2max~ during exercise tests to exhaustion. In a laboratory setting such tests commonly take place on treadmills or cycling ergometer. Fatigue prior to reaching the V̇O~2max~ results in underestimating its true value. Therefore most exercise protocols consist of a continuous or stepwise increase in load with an optimal duration of 8-12 minutes [@buchfuhrer1983 ; @yoon2007]. While these narrow limits have been challenged, considerably shorter or longer protocols should not be used [@midgley2008]. Despite protocol type and protocol duration, factors such as motivation [@midgley2018], exercise modality [@myers1991] and biological variability [@katch1982] bias the V̇O~2max~ determination.

Early research on V̇O~2max~ used bags to collect expired air for later analysis --- the so called Douglas bag method [@shephard2017]. Today's modern metabolic carts allow for simultaneous collection and analysis of gases, using either mixing chambers (which often sample data at fixed time intervals, e.g., 15 seconds) or breath-by-breath methods [@macfarlane2001]. While mixing chambers are regarded as more reliable [@beijst2013], measuring breath-by-breath generates data with a higher temporal resolution [@roecker2005]. The measuring method and the metabolic cart model can influence the measured oxygen uptake --- and as such the V̇O~2max~ [@winkert2021; @miles1994]. But even the same oxygen uptake data generated during an exercise test may result in varying outcomes when processed differently.

Whether an observed peak in oxygen uptake corresponds to the true maximum has been extensively discussed in the past decades [@taylor1955; @howley1995; @poole2008; @poole2017; @green2018]. To identify a true maximum, researchers commonly evaluate a set of parameters measured during ramp tests --- the so called 'V̇O~2max~ criteria' [@howley1995]. If these criteria are (partly) not fulfilled, researchers are advised to speak of a peak oxygen uptake instead of V̇O~2max~ [@howley1995]. In this thesis I will not distinguish between peak and maximum oxygen uptake, as the criteria for V̇O~2max~ (e.g. the primary criterion of a plateau in oxygen uptake or the secondary criterion of the maximum respiratory quotient) do heavily rely on the data processing strategy used [@astorino2009]. Thus, I speak of V̇O~2max~ as the maximum oxygen uptake measured during an appropriate exercise test regardless of any V̇O~2max~ criteria.

## Previous Research

Measured breath-by-breath data is noisy (see grey points in @fig-example). Both the biological variability of breathing patterns and the measurement error (as well as irregular breaths such as coughs and swallowing) lead to a highly fluctuating raw oxygen uptake. For interpretation the raw data requires some form of processing.

Different data processing strategies influence measured parameters of gas exchange. As soon as when the first automated systems for gas exchange measurement were available, Matthews et al. @matthews1987 reported that different processing strategies lead to different V̇O~2max~ values. Since then, many researchers investigated the influence of different data averaging intervals on V̇O~2max~, unsurprisingly showing that shorter calculation intervals lead to higher V̇O~2max~ values [@johnson1998; @astorino2009; @sell2021]. @midgley2007 Midgley et al. found differences between data processing strategies for V̇O~2max~, but no difference in the reliability of these, infering that no optimal strategy exists, but that consistency of strategies is key when comparing outcomes. Based on own data in sedentary and moderately trained individuals, @martin-rincon2019 Martin-Rincon et al. provided linear-log equations to compare mean V̇O~2max~ values derived from processing strategies with differing averaging interval lengths. There is inconclusive evidence on whether the influence of different processing strategies interacts with different exercise protocols [@hill2003; @scheadler2017].

```{r}
#| echo: false
#| fig-cap: "Example of raw breath-by-breath data processed by different strategies during the final minutes of a ramp test to exhaustion. Grey points display oxygen uptake from the single breaths. The moving averages were calculated over 30 breaths and seconds, respectively. The binned average was calculated over 30-second intervals with the mean values aligned to the center of each interval. For digital filtering a forward-backward (zero phase) low-pass Butterworth filter with the parameters 0.04 Hz (cut-off frequency) and 3 (order) for each filter was applied. For more details on the used data processing strategies, see the methods section. Data Source: [@quittmann_unpubl_b]"
#| fig-scap: "Example of raw breath-by-breath data processed by different strategies during the final minutes of a ramp test to exhaustion."
#| label: fig-example
#| out-width: '100%'

knitr::include_graphics("../plots/example.png")
```

Differences in V̇O~2max~ due to data processing can cause serious implications in practice. They hinder the assessment of longitudinal data from athletes that participated in diagnostics with differing data processing or the comparability of data from studies within meta-analysis [@martin-rincon2019]. Data processing strategies directly affect the occurrence of a plateau in oxygen uptake, the primary criterion for V̇O~2max~ [@astorino2009]. Crucially, in situations where individuals are classified by their V̇O~2max~ --- for example when describing the training status of a study population [@pauw2013a; @decroix2016] or evaluating patients for a heart transplantation [@mancini1991] --- differing processing strategies can lead to misclassifications [@johnson1998].

Calls to standardize data processing strategies of gas exchange data are frequent [@scheadler2017; @myers1990; @hill2003; @johnson1998]. Howley et al. @howley1995 argued to use longer calculation intervals (60 seconds), as shorter intervals may introduce bias towards extreme data values and thus could systematically overestimate true V̇O~2max~. Based on synthetic and experimental data, Robergs & Burnett @robergs2003 opposed this conclusion, stating that shorter calculation intervals are less erroneous. They further argued to use breath-based moving averages instead of time averages. In terms of data processing, the current ATS/ACCP guidelines for cardiopulmonary exercise testing remain vague, stating that *"30 to 60-second intervals for averaging data are recommended, although 20-seconds intervals may be acceptable*" [@ats/accp2003]. In a classic opinion piece, Robergs et al. @robergs2010 argued for using digital filtering (i.e. a low-pass Butterworth filter) for processing gas exchange data, a strategy initially introduced for gas exchange measures by Weir et al. @weir2004. However, @robergs2010 Robergs et al. acknowledged the lack of accessible software implementing such procedures.

In absence of one commonly accepted method, data processing varies among the literature. In light of the influence on outcome variables, many articles highlighted the need to report processing strategies in research articles [@myers1990; @astorino2009; @scheadler2017; @johnson1998]. @midgley2007 Midgley et al. were the first to evaluate reported data processing strategies for breath-by-breath analyses in selected journals. They found that almost all studies reporting their methods choose binned time averaging, with only 1 in 117 using a moving time and a moving breath average, respectively. One third of the studies did not describe their processing method at all. While providing interesting first insights into reporting and processing practices, the search by Midgley et al. @midgley2007 was not systematic and its methods were not described in detail. Robergs et al. @robergs2010 wrote that to investigate the current state of data processing strategies, two possible approaches are *"(i) a summary of published research, and (ii) a survey circulated via the Internet to as many exercise physiologists as possible"*. They provided the latter one with a total of 75 respondents, who reported a great variety in data processing strategies. Most researchers reported the use of binned time averages over 30 or 60 seconds. Astonishingly, about half of the respondents admitted that their data processing strategy was rather chosen due to subjective influences as opposed to objective reasoning. Practices in reporting and processing may have changed more than a decade after the publication of such numbers.

## Aim

In this thesis I will review the usage and reporting of different data processing strategies in the scientific literature and investigate their influence on the V̇O~2max~. To date, the practices of data processing to determine V̇O~2max~ in the actual scientific work remain largely unknown. Previous research on this topic has only performed unsystematic searches [@midgley2007] or non-representative surveys [@robergs2010]. Moreover, recent recommendation [@robergs2010] as well as advances in measurement devices and analysis software may have changed processing practices in the past 15 years.

Selected data processing strategies have been extensively compared in the literature. Due to the absence of a systematic mapping of current practices, these works lacked the reasoning of which strategies to compare. Many studies compared different averaging intervals, but not averaging types (e.g. moving breath vs. binned time) [@astorino2009; @midgley2007]. @martin-rincon2019 Martin-Rincon et al. provided formulas for comparing data processing strategies by investigating a data set of sedentary or recreational trained athletes, using two different metabolic carts. Therefore in this work motivation and measurement devices may have interacted with the influence of processing strategies.

Differences in V̇O~2max~ due to data processing strategies may have serious implication for individuals [@johnson1998], yet almost all comparisons only report mean differences between strategies. No research has yet compared a variety of systematically derived strategies among a group of well-trained individuals using a standardized measurement set-up.

This thesis will first map current practices of data processing for V̇O~2max~ determination by the means of a systematic scoping review of current scientific literature. Secondly, I will compare different data processing strategies in relation to the most commonly applied on a set of 72 standardized treadmill tests in well-trained individuals. The results will help to compare V̇O~2max~ data derived from different processing methods among studies and in individuals. The review allows to assess the implementation of current data processing recommendations and to find malpractices in reporting. The results build a basis for providing new recommendations for data processing and its reporting for determining the V̇O~2max.~