### 2022-05-31 transparent_changes.Rmd
Update transparent changes

---
title: ''
output: pdf_document
bibliography: references.bib
---

This document includes all deviations and modifications of code and methods in the final project compared to its [preregistration](https://osf.io/3am4s).

### Major Changes

#### Number of exercise tests for comparison

Due to a miscalculation, the preregistration provided an incorrect number of exercise test (n = 76). The correct number of exercise tests is n = 72, with 44 from @quittmann_unpubl_a (one test only partly included in the original work due to missing other data) and 28 from @quittmann_unpubl_b (three tests only partly included in the original work due to missing other data).

#### Additional variables extracted from included research

`outcome`: Which type of outcome V̇O~2max~ is. Either `primary`, `secondary` or `other`

`source`: Which source is provided for the data processing method used.

### Minor Changes

#### Code changes for automated article filtering and screening preparation

-   advanced detection of missing DOIs: `is.na(merge_data$doi) | (merge_data$doi == "")` instead of `is.na(merge_data$doi)`.
-   Improved function to retrieve missing PubMed abstracts: Handles case when input (PMID) is missing (`if (is.na(pmid)) return(NA)`).
-   save/load of the sampling results as an .Rda file to reduce computation time when working on parts of the workflow.

#### Unblinding of single abstracts

-   Manual retrieval of abstracts for articles, as these were neither present in the search result data, nor could be automatically scraped. Abstracts were saved and imported as .txt files. Temporary unblinding only applied to the primary researcher. This concerns the abstracts with the sampling id (sid): 50, 238, 288, 416, 488, 490, 500
-   Manual retrieval of abstracts for articles, as the automatically collected abstract contained html-tags that could not be removed for later abstract plots. Abstracts were saved and imported as .txt files. Temporary unblinding only applied to the primary researcher. This concerns the abstracts with the sampling id (sid): 344, 356
-   Unblinding during screening to assess the implications of title given in squared brackets. This concerns the abstracts with the sampling id (sid): 262, 303
-   Consulting of online abstract due to incomplete abstract plot. This concerns the abstract with the sampling id (sid): 275

#### Minor Modification of exclusion criteria

Changes are in italics:

`'r'`: Is the article no original research *(i.e. no primary analysis of experimental data)* ? (if yes, indicate 'r'; if no, continue)

`'t'`: Was no full-text ~~available~~ *accessible* for the corresponding article? (if yes, indicate 't', if no continue)

#### Minor screening error for two articles

For two articles (sampling ids: 194, 282) I only realized during data extraction that they matched the exclusion criteria (\`c\`: no continuous measurement of oxygen uptake). In agreement of both screeners, the screening data was retrospectively changed.